---
title: "Loading and Summarizing Data"
output: 
  learnr::tutorial:
   progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
gradethis::gradethis_setup()
```

## Introduction

In this tutorial, you will load data into R, and Summarize that data. You will use the following R functions: `read.csv`, `url`, `names`, `head`, `summary`, `attach`, `table`, `mean`, `median`, `var`, `sd`, `min`, `max`, `range`, `quantile`, `cor`. 

The following video is provides a nice overview of loading and summarizing data. Watch the whole video, and **do not worry about getting every detail.** This tutorial will walk you through each step.

![](https://youtu.be/ACWuV16tdhY)

## Loading Data

### How to load a data set

![](https://youtu.be/ACWuV16tdhY)

### A closer look at the code

For this tutorial, we will be using a data set that contains information on 78 people using one of three diets. This data is from the University of Sheffield's *Datasets for Teaching*, see https://www.sheffield.ac.uk/mash/statistics/datasets.

To load this data, we will use the `read.csv()` function along with the `url()` function.

```{r LoadData, echo=T}
diet <- read.csv(url("https://www.sheffield.ac.uk/polopoly_fs/1.886047!/file/Diet_R.csv"))

```

`<-` is the assignment operator. It is used to store something “into” a an object or variable. For now, think of an object as a box, and `<-` puts stuff in the box. So in the code fragment above, we are storing the data into an object called `diet`.

`diet` is a special object called a data frame. You can think of a data frame as one page of an workbook from a spreadsheet program.

We need to use the `url()` function because we are accessing the data file from the internet. If you access data localy, you do not need the `url()` function.

### Now you try

Load the diet data set into a data frame named diet.

```{r  LoadDataEX, exercise=TRUE}

```

```{r  LoadDataEX-solution}
diet <- read.csv(url("https://www.sheffield.ac.uk/polopoly_fs/1.886047!/file/Diet_R.csv"))
```

```{r  LoadDataEX-check}
grade_code("")
```

## Check the Data

### How to preforem an initial examination of the data set

![](https://youtu.be/U4-RnTW5dfw)

### A closer look at the code

It is always a good idea to look at a data set when you first load it. We need to check that the data loaded as we expected and determine the types and names of the variables in the data set. We can do this with the `head()`, `tail()` and `dim()` functions.

First, let's look at the first few rows of the data frame. We do this with the `head()` function

```{r FirstLookHead, echo=T}
head(diet)
```

Next, use the `tail()` function to look at the last few rows of the data frame.

```{r FirstLookTail, echo=T}
tail(diet)
```

Next, use the `dim()` function to see how many observation (rows) and variables (columns) the data frame has.

```{r FirstLookDim, echo=T}
dim(diet)
```

So we can see that the data set has 78 observation (rows) and 7 variables (columns).

### Now you try

Display the first and last rows of the diet data frame, and then display the dimensions of the diet data frame.

```{r LoadDataSetup, echo=F}
diet <- read.csv(url("https://www.sheffield.ac.uk/polopoly_fs/1.886047!/file/Diet_R.csv"))

```

```{r  FirstLookEX, exercise=TRUE, exercise.lines=9, exercise.setup = "LoadDataSetup"}
# First rows of the data frame


# Last rows of the data frame


# Dimensions of the data frame


```

```{r  FirstLookEX-solution}
# First rows of the data frame
head(diet)

# Last rows of the data frame
tail(diet)

# Dimensions of the data frame
dim(diet)
```

```{r  FirstLookEX-check}
grade_code("")
```

## Frequency Tables

### Summerizing catagorical data with frequiency tables

![](https://www.youtube.com/watch?v=ACWuV16tdhY&t=52s)

### A closer look at the code

A data frame, such as `diet`, contains variables that can be accessed by using a `$`. For example, `diet$Age` returns the observations of the variable Age. Age is a numeric variable, but gender is different. gender is a categorical variable that indicates the biological sex of the subject. 

The `table()` function creates counts of categorical data. For example, gender, takes the value 1 if the subject is male and 0 if the subject is female. The command

```{r, echo=T}
table(diet$gender)
```

creates a frequency table of smokers, and

```{r, echo=T}
table(diet$gender)/length(diet$gender)
```

creates a proportion table. Note `length()` returns the number of observations in gender. 

Because gender takes the value 1 if the subject is male and 0 if the subject is female, the table produced has column heads of 0 and 1. 

### Now you try

Create a table that count the frequency and proportion of subjects on each diet type by tabulating the variable Diet, i.e. `diet$Diet`. **Warning** R is case sensitive.

```{r TableEX, exercise=TRUE, exercise.lines=6, exercise.setup = "LoadDataSetup"}
# Frequency Table


# Proportion Table


```

```{r  TableEX-solution}
# Frequency Table
table(diet$Diet)

# Proportion Table
table(diet$Diet)/length(diet$Diet)

```

```{r  TableEX-check}
grade_code("")
```


## Contingency Tables

![](https://www.youtube.com/watch?v=ACWuV16tdhY&t=111s)

## Summary Statistics



![](https://www.youtube.com/watch?v=ACWuV16tdhY&t=129s)

## Correlation

![](https://www.youtube.com/watch?v=ACWuV16tdhY&t=293s)

## Min, Max and Range

## Quantiles






### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

